### 2021-12-12 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@5e6aba817564f072386d39b379e62faa00c327eb 🚀

\begin{itemize}
\tightlist
\item
  example on diagnostic test: rare events are hard to detect even with
  really good models
\item
  summary of model challenges for networks
\item
  Strydom et al. (2021) importance of drawing on traits + validation is
  challenging + comparing across space
\end{itemize}

@@

@@

In this section, we use a collection of XX empirical networks

\hypertarget{using-t-sbn-and-rdpg-to-simulate-arbitrarily-imperfect-classifier}{%
\subsection{Using t-SBN and RDPG to simulate arbitrarily imperfect
classifier}\label{using-t-sbn-and-rdpg-to-simulate-arbitrarily-imperfect-classifier}}

use t-SVD to only use some information

cumulative eigenvalues up to rank n give amount of information recovered

can be thresholded

how do values react to using fewer ranks

@@

\ldots{}

@@

@@

@@

@@

>>>>

\begin{itemize}
\tightlist
\item
  example on diagnostic test: rare events are hard to detect even with
  really good models
\item
  summary of model challenges for networks
\item
  Strydom, Catchen, et al. (2021) importance of drawing on traits +
  validation is challenging + comparing across space
\end{itemize}

@@

ROC-AUC

PR-AUC

@@

In this section, we use a collection of 22 empirical networks (Thompson
\& Townsend, 2004; Thompson \& Townsend, 1999) from streams in NEw
Zealand; these networks were previously analyzed by Pomeranz et al.
(2018), who found that a variety of rule-based models tended to increase
the number of interactions. Our goal here is not to infer which
interactions are missing, but rather to estimate (i) how much precise
knowledge of the network influences the measures of classification
performance, and (ii) whether an increase in performance also results in
more ecologically realistic networks. In order to do so, we use a
trucated Singular Value Decomposition (t-SVD) coupled to a Random Dot
Product Graph (RDPG; \emph{e.g.} Dalla Riva \& Stouffer, 2015) at
increasingly high dimensions to reconstruct an approximation of the
network. Recent work by Poisot et al. (2021) and Strydom, Bouskila, et
al. (2021) show that t-SVD and RDPG are able to infer biotic
interactions in, respectively, host-virus and predator-prey systems. The
t-SVD/RDPG approach is covered in detail by Strydom, Bouskila, et al.
(2021); by approximating the network at different ranks, where rank 1 is
the coarsest, we can generate a matrix that can be thresholded and
compared to the original network. The variance explained at each rank is
given by the relative sum of the eigenvalues of the SVD up to the chosen
rank, and measures how much information about the network we can capture
(Strydom, Dalla Riva, et al., 2021).

@@

By changing the rank of approximation from 1 (variance explained is
minimized) to 45 (variance explained is close to 1), we show in
fig.~\ref{fig:svdperf} that although all measures of classification
performance increase the more we know about the network, MCC is
initially lower compared to informedness; similarly, ROC-AUC is
consistently higher compared to PR-AUC.

\begin{figure}
\hypertarget{fig:svdperf}{%
\centering
\includegraphics{figures/svd_perf.png}
\caption{TODO}\label{fig:svdperf}
}
\end{figure}

\hypertarget{effect-of-classifier-performance-on-network-properties}{%
\subsection{Effect of classifier performance on network
properties}\label{effect-of-classifier-performance-on-network-properties}}

\begin{figure}
\centering
\includegraphics{figures/svd_ecol.png}
\caption{TODO}
\end{figure}

connectance, spectral radius (Staniczenko et al., 2013), entropy (Stock
et al., 2021)

@@

\leavevmode\hypertarget{ref-DallaRiva2015ExpEvo}{}%
Dalla Riva, G. V., \& Stouffer, D. B. (2015). Exploring the evolutionary
signature of food webs' backbones using functional traits. \emph{Oikos},
\emph{125}(4), 446--456. \url{https://doi.org/10.1111/oik.02305}

@@

\leavevmode\hypertarget{ref-Poisot2021ImpMam}{}%
Poisot, T., Ouellet, M.-A., Mollentze, N., Farrell, M. J., Becker, D.
J., Albery, G. F., Gibb, R. J., Seifert, S. N., \& Carlson, C. J. (2021,
May 31). \emph{Imputing the mammalian virome with linear filtering and
singular value decomposition}. \url{http://arxiv.org/abs/2105.14973}

\leavevmode\hypertarget{ref-Pomeranz2018InfPre}{}%
Pomeranz, J. P., Thompson, R. M., Poisot, T., \& Harding, J. S. (2018).
Inferring predator-prey interactions in food webs. \emph{Methods in
Ecology and Evolution}, \emph{0}.
\url{https://doi.org/10.1111/2041-210X.13125}

@@

\leavevmode\hypertarget{ref-Staniczenko2013GhoNes}{}%
Staniczenko, P. P. A., Kopp, J. C., \& Allesina, S. (2013). The ghost of
nestedness in ecological networks. \emph{Nature Communications},
\emph{4}(1), 1391. \url{https://doi.org/10.1038/ncomms2422}

\leavevmode\hypertarget{ref-Stock2021DisInf}{}%
Stock, M., Hoebeke, L., \& De Baets, B. (2021). Disentangling the
Information in Species Interaction Networks. \emph{Entropy},
\emph{23}(6, 6), 703. \url{https://doi.org/10.3390/e23060703}

\leavevmode\hypertarget{ref-Strydom2021FooWeb}{}%
Strydom, T., Bouskila, S., Banville, F., Barros, C., Caron, D., Farrell,
M. J., Fortin, M.-J., Hemming, V., Mercier, B., Pollock, L., Runghen,
R., Riva, G. V. D., \& Poisot, T. (2021). \emph{Food web reconstruction
through phylogenetic transfer of low-rank network representation}.
\url{https://doi.org/10.32942/osf.io/y7sdz}

@@

\leavevmode\hypertarget{ref-Strydom2021SvdEnt}{}%
Strydom, T., Dalla Riva, G. V., \& Poisot, T. (2021). SVD Entropy
Reveals the High Complexity of Ecological Networks. \emph{Frontiers in
Ecology and Evolution}, \emph{9}.
\url{https://doi.org/10.3389/fevo.2021.623141}

\leavevmode\hypertarget{ref-Thompson2004LanUse}{}%
Thompson, R. M., \& Townsend, C. R. (2004). Land‐use influences on New
Zealand stream communities: Effects on species composition, functional
organisation, and food‐web structure. \emph{New Zealand Journal of
Marine and Freshwater Research}, \emph{38}(4), 595--608.
\url{https://doi.org/10.1080/00288330.2004.9517265}

\leavevmode\hypertarget{ref-Thompson1999EffSea}{}%
Thompson, R. M., \& Townsend, C. R. (1999). The Effect of Seasonal
Variation on the Community Structure and Food-Web Attributes of Two
Streams: Implications for Food-Web Science. \emph{Oikos}, \emph{87}(1),
75--88. \url{https://doi.org/10.2307/3546998}
